---
title: "team"
author: "Xinyi Zhu, Han Sun"
date: "3/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Part 1 Cleaning Data Frame

```{r}
library(ggplot2)
library(classdata)
library(dplyr)
library(tidyr)
library(readxl)
# import defense sheet
defense <- read_excel('/Users/jessie/Desktop/ds202/cyclonesFootball2019.xlsx', sheet='Defensive')
str(defense)
# import offense sheet
offense <- read_excel('/Users/jessie/Desktop/ds202/cyclonesFootball2019.xlsx', sheet='Offensive')
str(offense)
# import biography sheet
biography <- read_excel('/Users/jessie/Desktop/ds202/cyclonesFootball2019.xlsx', sheet='Biography')
str(biography)
```

The player names and opponent names are factors.

```{r}
defense$Name <- factor(defense$Name)
defense$Opponent_Opponent <- factor(defense$Opponent_Opponent)
str(defense)

offense$Name <- factor(offense$Name)
offense$Opponent_Opponent <- factor(offense$Opponent_Opponent)
str(offense)
```

All offensive and defensive statistics are numerical. The Weight column in biogrpahy is also numeric.

```{r}

defClean <- defense %>%
  mutate(Tackles_Solo = as.numeric(Tackles_Solo),
         Tackles_ASST = as.numeric(Tackles_ASST),
         Tackles_TFL = as.numeric(Tackles_TFL),
         Tackles_Sack = as.numeric(Tackles_Sack),
         Turnover_FF = as.numeric(Turnover_FF),
         Turnover_FR = as.numeric(Turnover_FR),
         Turnover_INT = as.numeric(Turnover_INT),
         Pass_QBH = as.numeric(Pass_QBH),
         Pass_PB = as.numeric(Pass_PB))
str(defClean)

offense$Name <- factor(offense$Name)
offense$Opponent_Opponent <- factor(offense$Opponent_Opponent)
names(offense)[9] <- "Passing_CMP_ATT"
offClean <- offense %>%
  mutate(Rushing_ATT = as.numeric(Rushing_ATT),
         Rushing_YDS = as.numeric(Rushing_YDS),
         Rushing_TD = as.numeric(Rushing_TD),
         Receiving_REC = as.numeric(Receiving_REC),
         Receiving_YDS = as.numeric(Receiving_YDS),
         Receiving_TD = as.numeric(Receiving_TD),
         Passing_CMP_ATT = as.numeric(Passing_CMP_ATT),
         Passing_YDS = as.numeric(Passing_YDS),
         Passing_TD = as.numeric(Passing_TD),
         Passing_INT = as.numeric(Passing_INT))
names(offense)[9] <- "Passing_CMP_ATT"
offClean[is.na(offClean)] = 0
str(offClean)

bioClean <- biography %>%
  mutate(Weight = as.numeric(Weight))
str(bioClean)

```

Change the Height column in biogrpahy into numeric. 

```{r}
bioClean<-biography %>% separate(Height, into = c("feet", "inch"))
bioClean <- bioClean %>%
  mutate(feet = as.numeric(feet),
         inch = as.numeric(inch))
bioClean$Height <- bioClean$feet+bioClean$inch/12
bioClean
```

By cleaning the three data frames in R, at the end of this step you should have the following three clean data frames: defClean, offClean, and bioClean. Print the structures of these data frames.

```{r}
defClean
offClean
bioClean
```
 
Part two: Tidying 

Reformat the defClean data frame into a tidy format using pivot_longer. The type of statistic (Tackles_Solo, Tackles_ASST, etc) is added as a new key column named Statistic.

```{r}
defClean1 <- defClean %>% pivot_longer(3:11, names_to = 'Statistic', values_to = 'score')
head(defClean1)
```

Part three: Joining data frames

1. Find where do the offensive players come from by joining the offensive data and player biography data appropriately.

```{r}
left<- left_join(offClean, bioClean, by = "Name")
left
```


2. Is there a relationship between the weight of a player and the receiving yards (Receiving_YDS)?

```{r,fig.width=18, fig.height=20}
ggplot(data =left, aes(x = Weight, y = Receiving_YDS)) +
  geom_point()
```
 
The histogram does not show obvious relationship between weight and Receiving_YDS. However, But in weight 183, it's more evenly distributed in Receiving_YDS.

```{r}
inner<- inner_join(offClean, bioClean, by = "Name")
inner
pb <-inner[inner$Name %in% c("Purdy, Brock"), ]

pb1 <- pb %>%
  select("Rushing_ATT", "Rushing_YDS",  "Rushing_TD", "Passing_YDS", "Passing_TD", "Passing_INT")
pb1

hist(pb1$Rushing_ATT)
hist(pb1$Rushing_YDS)
hist(pb1$Rushing_TD)
hist(pb1$Passing_YDS)
hist(pb1$Passing_TD)
hist(pb1$Passing_INT)
```

