---
title: "team"
author: "Xinyi Zhu, Han Sun"
date: "3/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Part one: Cleaning Data Frame

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(readxl)
# import defense sheet
defense <- read_excel('cyclonesFootball2019.xlsx', sheet='Defensive')
# import offense sheet
offense <- read_excel('cyclonesFootball2019.xlsx', sheet='Offensive')
# import biography sheet
biography <- read_excel('cyclonesFootball2019.xlsx', sheet='Biography')
```

1. The player names and opponent names are factors.

```{r}
defense$Name <- as.factor(defense$Name)
defense$Opponent_Opponent <- as.factor(defense$Opponent_Opponent)

offense$Name <- as.factor(offense$Name)
offense$Opponent_Opponent <- as.factor(offense$Opponent_Opponent)

biography$Name <- as.factor(biography$Name)
```

2. All offensive and defensive statistics are numerical. The Weight column in biogrpahy is also numeric.

```{r}
defClean <- defense %>%
  mutate(Tackles_Solo = as.numeric(Tackles_Solo),
         Tackles_ASST = as.numeric(Tackles_ASST),
         Tackles_TFL = as.numeric(Tackles_TFL),
         Tackles_Sack = as.numeric(Tackles_Sack),
         Turnover_FF = as.numeric(Turnover_FF),
         Turnover_FR = as.numeric(Turnover_FR),
         Turnover_INT = as.numeric(Turnover_INT),
         Pass_QBH = as.numeric(Pass_QBH),
         Pass_PB = as.numeric(Pass_PB))

names(offense)[9] <- "Passing_CMP_ATT"
offClean <- offense %>%
  mutate(Rushing_ATT = as.numeric(Rushing_ATT),
         Rushing_YDS = as.numeric(Rushing_YDS),
         Rushing_TD = as.numeric(Rushing_TD),
         Receiving_REC = as.numeric(Receiving_REC),
         Receiving_YDS = as.numeric(Receiving_YDS),
         Receiving_TD = as.numeric(Receiving_TD),
         Passing_CMP_ATT = as.numeric(Passing_CMP_ATT),
         Passing_YDS = as.numeric(Passing_YDS),
         Passing_TD = as.numeric(Passing_TD),
         Passing_INT = as.numeric(Passing_INT))
offClean[is.na(offClean)] = 0

bioClean <- biography %>%
  mutate(Weight = as.numeric(Weight))
```

3. Change the Height column in biogrpahy into numeric. 

```{r}
bioClean <- bioClean %>% 
  separate('Height', into = c("Feet", "Inch")) %>%
  mutate_at('Feet', as.numeric)%>%
  mutate_at('Inch', as.numeric)
bioClean$Height <- bioClean$Inch + bioClean$Feet * 12
head(bioClean)
```

By cleaning the three data frames in R, at the end of this step you should have the following three clean data frames: defClean, offClean, and bioClean. Print the structures of these data frames.

```{r}
str(defClean)
str(offClean)
str(bioClean)
```
 
### Part two: Tidying (1)

1. Reformat the defClean data frame into a tidy format using pivot_longer. The type of statistic (Tackles_Solo, Tackles_ASST, etc) is added as a new key column named Statistic.

```{r}
defClean <- defClean %>% 
  pivot_longer(Tackles_Solo:Pass_PB, names_to = 'Statistic')
head(defClean)
```

2. Create a histogram for each defensive statistic. Use facet_wrap to draw all histograms in a single plot. What defensive skills are rare?

```{r}
defClean %>% 
  ggplot(aes(value)) + geom_histogram(binwidth = 0.5) + facet_wrap(~Statistic)
```
From the defensive histograms, the Tackles_Sack, Turnover_FF, Turnover_FR, Turnover_INT skills are rare.

3. Did ISU have better defense against West Virginia or Kansas State? Answer this question by creating a scatterplot with x-axis being the number of solo tackles (by each player) against West Virginia, and y-axis being that against Kansas State. A large number of solo tackles indicates good defense.

```{r}
defClean %>%
  filter(Statistic == "Tackles_Solo", Opponent_Opponent %in% c("West Virginia", "Kansas State")) %>%
  spread(Opponent_Opponent,value) %>%
  ggplot(aes(x = `West Virginia`, y = `Kansas State`)) + geom_count() + geom_abline(slope = -1, intercept = 5, colour = "orange")
```

Base on the scatterplot, most points are below the line. It means, most individual players have a higher number of solo tackles against Kansas State than against West Virginia.

### Part two: Tidying (2)

1. Separate the Hometown column into two columns, namely the city and the state. Print the first few rows of these two columns. (Hint: look at the sep= argument of separate)

```{r}
bioClean <- bioClean %>%
  separate(Hometown, into = c('City', 'State'), sep = ",")
head(bioClean)
```

2. How many players are there from each state?

```{r}
table(bioClean$State)
```

### Part three: Joining data frames

1. Find where do the offensive players come from by joining the offensive data and player biography data appropriately.

```{r}
left <- left_join(offClean, bioClean, by = "Name")
leftns <- left %>%
   select("Name", "State")
unique(leftns)
```

You can looking for hometown column from the table then find where do the offensive players come from.

2. Is there a relationship between the weight of a player and the receiving yards (Receiving_YDS)?

```{r,fig.width=18, fig.height=20}
full <- full_join(offClean, bioClean, by = "Name")
ggplot(data =full, aes(x = Weight, y = Receiving_YDS)) +
  geom_point()
```
 
The histogram does not show obvious relationship between weight and Receiving_YDS. However, But in weight 183, it's more evenly distributed in Receiving_YDS.

3. How do the performance of quarterback Brock Purdy relate to the performance of the ISU team?

```{r,fig.width=12, fig.height=5}
inner <- inner_join(offClean, bioClean, by = "Name")
TeamISU <- inner %>%
  filter(State == " Iowa") %>%
  group_by(Opponent_Opponent) %>%
  summarize(Rushing_YDS = sum(Rushing_YDS, na.rm = T),
            Rushing_ATT = sum(Rushing_ATT, na.rm = T),
            Rushing_TD = sum(Rushing_TD, na.rm = T),
            Receiving_REC = sum(Receiving_REC, na.rm = T),
            Receiving_YDS = sum(Receiving_YDS, na.rm = T),
            Receiving_TD = sum(Receiving_TD, na.rm = T),
            Passing_CMP_ATT = sum(Passing_CMP_ATT, na.rm = T),
            Passing_YDS = sum(Passing_YDS, na.rm = T),
            Passing_TD = sum(Passing_TD, na.rm = T),
            Passing_INT = sum(Passing_INT, na.rm = T)) %>%
  mutate(ID = "Team ISU")
BP <- inner %>%
  filter(Name == "Purdy, Brock") %>%
  group_by(Opponent_Opponent) %>%
  summarize(Rushing_YDS = sum(Rushing_YDS, na.rm = T),
            Rushing_ATT = sum(Rushing_ATT, na.rm = T),
            Rushing_TD = sum(Rushing_TD, na.rm = T),
            Receiving_REC = sum(Receiving_REC, na.rm = T),
            Receiving_YDS = sum(Receiving_YDS, na.rm = T),
            Receiving_TD = sum(Receiving_TD, na.rm = T),
            Passing_CMP_ATT = sum(Passing_CMP_ATT, na.rm = T),
            Passing_YDS = sum(Passing_YDS, na.rm = T),
            Passing_TD = sum(Passing_TD, na.rm = T),
            Passing_INT = sum(Passing_INT, na.rm = T)) %>%
  mutate(ID = "Brock Purdy")
Sum <- full_join(TeamISU, BP, by = c("ID", "Opponent_Opponent", "Rushing_YDS", "Rushing_ATT", "Rushing_TD", "Receiving_REC", "Receiving_YDS", "Receiving_TD", "Passing_CMP_ATT", "Passing_YDS", "Passing_TD", "Passing_INT"))
ggplot(data = Sum, aes(x = Opponent_Opponent, y = Rushing_YDS, group = ID, col = ID)) + geom_line()
ggplot(data = Sum, aes(x = Opponent_Opponent, y = Rushing_ATT, group = ID, col = ID)) + geom_line()
ggplot(data = Sum, aes(x = Opponent_Opponent, y = Rushing_TD, group = ID, col = ID)) + geom_line()
ggplot(data = Sum, aes(x = Opponent_Opponent, y = Receiving_YDS, group = ID, col = ID)) + geom_line()
ggplot(data = Sum, aes(x = Opponent_Opponent, y = Receiving_REC, group = ID, col = ID)) + geom_line()
ggplot(data = Sum, aes(x = Opponent_Opponent, y = Passing_YDS, group = ID, col = ID)) + geom_line()
ggplot(data = Sum, aes(x = Opponent_Opponent, y = Passing_TD, group = ID, col = ID)) + geom_line()
ggplot(data = Sum, aes(x = Opponent_Opponent, y = Passing_INT, group = ID, col = ID)) + geom_line()
```

There is no obvious relationship betweenthe performance of quarterback Brock Purdy and the performance of the ISU team. Brock Purdy performances well in rushing skill. But he performances less than ISU team in receiving and passing skills.

4. Explore more. You may in addition explore the 2018 data here. Raise a question that can be answered only by looking at two or more data frames. Then answer that question.

Is there a relationship between the Height(in inch) of a player and the receiving yards (Receiving_YDS)?
```{r}
ggplot(data =full, aes(x = Height, y = Receiving_YDS)) +
  geom_point()
```

The histogram does not show obvious relationship between height and Receiving_YDS. However, But in height 70, it's more evenly distributed in Receiving_YDS.